---
title: "REDCap File Migration Functions"
output: 
  rmarkdown::html_document:
    toc: TRUE
    toc_float: TRUE
    css: redcap_project_migration_app/edu.wustl.ohids.css
  rmarkdown::pdf_document: default
vignette: >
  %\VignetteIndexEntry{REDCap File Migration Functions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wucap)
```

# Background

[Washington University in Saint Louis](https://wustl.edu/) (WUSTL) maintains a legacy [REDCap](https://www.project-redcap.org/) instance frozen at v7.3.5 due to undocumented modification of the codebase to support home-grown features. Migration of active projects from legacy to current production and archival of inactive projects are desirable due to failing hardware.

Upgrading REDCap in place is best practice but was never possible due to the codebase modifications. The next best approach is export and import of a [CDISC](https://www.cdisc.org/) Operational Data Model (ODM) extensible markup language (XML) file, but this fails for some projects because current production is unable to import large XML files.

Therefore, we are reduced to migrating projects in parts either using the web or application programming interface (API). The web interface approach is tedious (especially when file uploads are present), and using the API is our best option.

This vignette showcases the functions dealing with the migration of file uploads from one REDCap project to another.

# Package Installation

The [{wucap}](https://github.com/the-mad-statter/wucap) package was written and hosted on Github to make distribution and use of the code as convenient as possible.

Before running the app code, the package must be installed. To install the package run the following command in a Databricks workbook R cell:

```
remotes::install_github("the-mad-statter/wucap")
```

# File Transfer Functions

## SSL Certificate Checking

If it is necessary to turn off the default SSL certificate validation, the following code may be used:

```
httr::set_config(httr::config(ssl_verifypeer = 0L, ssl_verifyhost = 0L))
```

## Endpoints

Make a note of source and destination project API endpoints.

```
redcap_endpoint_src <- wucap:::redcap_api_endpoints$prod$v7.3.5

redcap_endpoint_dst <- wucap:::redcap_api_endpoints$prod$latest
```

## Tokens

Gather authorization tokens.

```
redcap_token_src <- retrieve_credential_local("1784")$token

redcap_token_dst <- retrieve_credential_local("14464")$token
```

## Local Repository Location

```
repo <- "C:/Users/schuelke/Desktop/redcap_download_files"
```

## Download Files from the Source Project

```
# read source data dictionary to determine file field names
data_dictionary_src <- REDCapR::redcap_metadata_read(
  redcap_endpoint_src, 
  redcap_token_src
)$data

# read source project records to identify existing uploads
project_records_src <- REDCapR::redcap_read(
  redcap_uri = redcap_endpoint_src,
  token = redcap_token_src
)$data

# construct file (primary) keys object to identify existing uploads
file_keys_src <- redcap_read_file_keys(
  data_dictionary_src, 
  project_records_src,
  field_type = "both"
)

# download the files to the local repo
redcap_export_files(
  file_keys_src,
  repo,
  redcap_endpoint_src,
  redcap_token_src
)

# read the file index to check for missing downloads due to download errors
file_index <- redcap_read_file_index(repo)

# identify any missing downloads
missing_downloads <- file_setdiff(file_keys_src, file_index)

# if there are missing downloads, export them from the source project
redcap_export_files(
  missing_downloads,
  repo,
  redcap_endpoint_src,
  redcap_token_src
)
```

## Upload Files to the Destination Project

```
# upload the files from the local repo
redcap_import_files(
  file_index,
  repo,
  redcap_endpoint_dst,
  redcap_token_dst
)

# read the destination project dictionary to identify file field names
data_dictionary_dst <- REDCapR::redcap_metadata_read(
  redcap_endpoint_dst, 
  redcap_token_dst
)$data

# read the destination project records to identify existing uploads
project_records_dst <- REDCapR::redcap_read(
  redcap_uri = redcap_endpoint_dst,
  token = redcap_token_dst
)$data

# construct a file (primary) key object to identify existing uploads
file_keys_dst <- redcap_read_file_keys(
  data_dictionary_dst, 
  project_records_dst,
  field_type = "both"
)

# check for missing uploads
missing_uploads <- file_setdiff(file_index, file_keys_dst)

# if there are any missing uploads, import them into the destination project
redcap_import_files(
  missing_uploads,
  repo,
  redcap_endpoint_dst,
  redcap_token_dst
)
```
